---
title: "Norges 1000 Rikeste - 2013"
author: "Turkuler"
date: "7/7/2020"
output: html_document
---

For this exploratory analysis, I used a newspaper article on Dagens Næringsliv on the page with below url, which lists 1000 (actually 997 but I saw it after I downloaded the table) richest people in Norway in 2013 according to previous year tax documents

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The libraries I need are htmltab for pulling the table from an html page, ggplot2 and pppubr for visualization and dplyr for data modification:

```{r libraries, message=FALSE}
library(htmltab)
library(ggplot2)
library(ggpubr)
library(dplyr)
```

I started with downloading html table to R and making some modifications to replace the column name of birth year (Født) with a Norwegian letter to all English characters and to numeric, commas to dots:
```{r loading}
url <- "https://www.dn.no/her-er-norges-1000-rikeste/1-1-2009497"
rikeste1000 <- htmltab(doc = url, which = "//th[text() = 'Navn']/ancestor::table") 
rikeste1000 <- rename(rikeste1000, Fodt = Født) %>%
               mutate(Fodt = as.numeric(Fodt),
                      Inntekt = as.numeric(gsub(",",".", Inntekt)), 
                      Skatt = as.numeric(gsub(",",".", Skatt)),
                      Formue = as.numeric(gsub(",",".", Formue)),
                      Bosted = as.factor(Bosted))
```

Checking the lookabouts of table:
```{r exploration1}
head(rikeste1000)
summary(rikeste1000)
```

First, the distributions of tax (Skatt), income (Inntekt) and wealth (Formue):
```{r exploration3}
ggplot(rikeste1000) + geom_boxplot(mapping = aes(y = Skatt))
ggplot(rikeste1000) + geom_boxplot(mapping = aes(y = Inntekt))
ggplot(rikeste1000) + geom_boxplot(mapping = aes(y = Formue))
```
There are many outliers that generally falls in the higher values for the listed parameters. Without doing anything about these outliers, I checked whether there is any relationship between these parameters with using scatter plots. I also added a regression line with confidence intervals and Pearson correlation value. These plots show very high correlations between income and tax, and wealth and tax, and a low correlation between wealth and income. However, the correlations are obviously affected from the outlier points.

```{r exploration4, message=FALSE}
ggscatter(data = rikeste1000, x = "Inntekt", y = "Skatt", add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), 
          conf.int = TRUE) + stat_cor(method = "pearson")
ggscatter(data = rikeste1000, x = "Formue", y = "Skatt", add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), 
          conf.int = TRUE) + stat_cor(method = "pearson")
ggscatter(data = rikeste1000, x = "Inntekt", y = "Formue", add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), 
          conf.int = TRUE) + stat_cor(method = "pearson")
```

I removed outlier points/people by using interquartile range in a function named outliers.

```{r outlierRemoval1}
outliers <- function(x){
  x > quantile(x)[4] + 1.5*IQR(x) | x < quantile(x)[2] - 1.5*IQR(x)
}

attach(rikeste1000)
outlier_inntekt <- sum(outliers(Inntekt))
outlier_inntekt_index <- which(outliers(Inntekt))
outlier_skatt <- sum(outliers(Skatt))
outlier_skatt_index <- which(outliers(Skatt))
outlier_formue <- sum(outliers(Formue))
outlier_formue_index <- which(outliers(Formue))

rikeste1000_orem <- slice(rikeste1000, -c(outlier_inntekt_index,outlier_skatt_index, outlier_formue_index)) 
```

There are total of `r nrow(rikeste1000) - nrow(rikeste1000_orem)` people having an outlier value for any of these 3 parameters, which now filtered out in the new dataset. Also, some people have 0 as income value (`r sum(rikeste1000$income == 0)`) but I will not remove them now. The new scatter plots are below and now the correlations are a bit lower but they seem more consistent with data points. The correlation between wealth and income is much more lower than the initial one.

```{r exploration5, message=FALSE}
ggscatter(data = rikeste1000_orem, x = "Inntekt", y = "Skatt", add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), 
          conf.int = TRUE) + stat_cor(method = "pearson")
ggscatter(data = rikeste1000_orem, x = "Formue", y = "Skatt", add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), 
          conf.int = TRUE) + stat_cor(method = "pearson")
ggscatter(data = rikeste1000_orem, x = "Inntekt", y = "Formue", add = "reg.line", 
          add.params = list(color = "blue", fill = "lightgray"), 
          conf.int = TRUE) + stat_cor(method = "pearson")
```

What else I can do with this data? `r emo::ji("thinking")`. Maybe I can look into the birth year decades! I will use the original dataset for this part.

```{r}
rikeste1000 <- mutate(rikeste1000, FodtBin = cut(Fodt, breaks =seq(1920, 2000, 10), dig.lab=4)) # since I know minimum and maximum values from summary above
ggplot(rikeste1000, aes(x=FodtBin)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(angle=45, vjust = 0.5)) + # to give an angle to x-axis labels
  xlab("Birth year") +
  ylim(NA, 300) + # to set the maximum of y-axis to make count values more visible
  geom_text(stat='count', aes(label=..count..), vjust=-1) # to add count values on top of bars
```

It seems like half of the richest people were in their 50s-60s at the time of this publication.

The last thing I want to look here is the places that these people were coming. There are `r group_by(rikeste1000, Bosted) %>% summarise(Unique_Places = n_distinct(Bosted)) %>% nrow(.)` places on Bosted column. That's too much for me. I will try to group these places into a different administrative unit in Norway: Fylke. First, I need another table:

```{r}
url <- "https://www.kartverket.no/Kunnskap/Fakta-om-Norge/Fylker-og-kommuner/Tabell/"
fylkeKommune <- htmltab(doc = url, which = "//table[@id='ctl00_FullRegion_UcTableDataGridView_DataTableGridView']") %>% 
  rename(Bosted = "Kommunar (356)") %>% mutate(Bosted = toupper(Bosted))
newdf <- left_join(rikeste1000, fylkeKommune, by = "Bosted")
```

Now, I stuck here since there are `r sum(is.na(newdf$Fylkesnummer))` "Bosted"s that are smaller administrative units than I expected and I couldn't find a source to map them into "Fylke" or "Kommune". To be continued...




